### 2015-02-22 zoldatoff/tex/pattern discovery/lecture02.tex
Уточнены формулы в разделе "Mining colossal patterns"
* Добавлены формулы из статьи «Mining Colossal Frequent Patterns by
Core Pattern Fusion»
* мелкие правки по форматированию

\section{Lecture 2: Pattern Discovery Basic Concepts}
\subsection{Frequent Itemsets (Patterns)}

X = itemset

\begin{itemize}
\item \textbf{(absolute) support (count) of X:} Frequency or the number of occurrences of an itemset X
\item \textbf{(relative) support, s:} The fraction of transactions that contains X (i.e., the probability that a transaction contains X)
\item An itemset X is \textbf{frequent} if the support of X is no less than a $minsup$ threshold (denoted as $\sigma$): $sup(X) \geqslant \sigma$.
\end{itemize}

%--
\subsection{Association Rules}
Association rules: $X \to Y (s, c)$:
\begin{itemize}
\item \textbf{Support} (s): the probability that a transaction contains $X \cup Y$:
\begin{equation*}
\mathrm{sup}(X \to Y) = \mathrm{P}(X \cup Y)
\end{equation*}

\item \textbf{Confidence} (c): the conditional probability that a transaction containing X also contains Y: 
\begin{equation*}
c = \mathrm{P}(Y \mid X) = \frac{sup(X \cup Y)}{sup(X)}
\end{equation*}
\end{itemize}

%--
\subsection{Expressing Patterns in Compressed Form}
Solution 1: \textbf{Closed patterns:} \textit{A pattern (itemset) X is closed if X is frequent, and there exists no super-pattern $Y \supset X$, with the same support as X}.\\

Closed pattern is a lossless compression of frequent patterns.\\

Solution 2: \textbf{Max-patterns:} \textit{A pattern X is a max-pattern if X is frequent and there exists no frequent super-pattern $Y \supset X$}.\\

Max-pattern is a lossy compression!

%--
\subsection{Recommended readings}
\begin{itemize}
\item R. Agrawal, T. Imielinski, and A. Swami, <<Mining association rules between sets of items in large databases>>, in Proc. of SIGMOD'93
\item R. J. Bayardo, <<Efficiently mining long patterns from databases>>, in Proc. of SIGMOD'98
\item N. Pasquier, Y. Bastide, R. Taouil, and L. Lakhal, <<Discovering frequent closed itemsets
for association rules>>, in Proc. of ICDT'99
\item J. Han, H. Cheng, D. Xin, and X. Yan, <<Frequent Pattern Mining: Current Status and Future Directions>>, Data Mining and Knowledge Discovery, 15(1): 55-86, 2007
\end{itemize}