### 2015-02-24 zoldatoff/tex/pattern discovery/lecture07.tex
Sequential Pattern Mining
Начата новая глава Sequential Pattern Mining

\section{Sequential Pattern Mining}

\subsection{Sequential Pattern}
Sequence $\to$ Element $\to$ Item or Event (items within an element are unordered)

The Apriori property still holds: if a subsequence $s_1$ is infrequent, none of $s_1$’s super-sequences can be frequent.\\

Algorithms:
\begin{itemize}
\item Generalized Sequential Patterns: \textbf{GSP}
\item Vertical format-based mining: \textbf{SPADE}
\item Pattern-growth methods: \textbf{PrefixSpan}
\item Mining closed sequential patterns: \textbf{CloSpan}
\end{itemize}

%--
\subsection{GSP: Apriori- Based Sequential Pattern Mining}

\begin{algorithm}
\caption{GSP}
\begin{algorithmic}
\State k = 1
\Repeat
    \State find length=k frequent sequences
    \State Apriori: remove candidates with sup < min\_sup
    \State length=k frequent sequences $\Rightarrow$ length=(k+1) candidate sequences
    \State k = k + 1
\Until{no frequent sequences or candidates}
\end{algorithmic}
\end{algorithm}

%--