### 2015-03-28 zoldatoff/tex/text retrieval/lecture01.tex
Лекции по курсу Text Retrieval
Лекции с 1-й по 5-ю

\section{Natural Language Content Analysis} 
NLP = Natural Language Processing

\subsection{An Example of NLP}
\begin{figure}[H]
    \centering
    \includegraphics[width=\linewidth]{NLP.png}
\end{figure}

\subsection{The State of the Art}
\begin{figure}[H]
    \centering
    \includegraphics[width=\linewidth]{NLP_state.png}
\end{figure}

\subsection{Recommended reading}
\begin{itemize}
\item Chris Manning and Hinrich Sch{\"u}tze, <<Foundations of Statistical Natural Language Processing>>, MIT Press. Cambridge, MA: May 1999.
\end{itemize}