### 2015-03-29 zoldatoff/tex/text retrieval/lecture07.tex
Лекции по курсу Text Retrieval
Лекции 8 и 9

\section{Vector Space Retrieval Model: Improved Instantiation}

Improved VSM:
\begin{itemize}
\item Dimension = word
\item Vector = TF-IDF weight vector
\item Similarity = dot product
\end{itemize}

\subsection{Improved VSM with Term Frequency (TF) Weighting}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{VSM_TF.png}
\end{figure}

\subsection{IDF Weighting: Penalizing Popular Terms}
IDF — inverse document frequency
\begin{figure}[H]
    \centering
    \includegraphics[width=0.65\linewidth]{IDF.png}
\end{figure}

\subsection{Adding Inverse Document Frequency (IDF)}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.65\linewidth]{VSM_IDF.png}
\end{figure}