### 2015-04-19 zoldatoff/tex/text retrieval/lecture12.tex

\newpage
\section{Course Summary}

\subsection{Recommended reading}
\begin{itemize}
\item \href{http://www.morganclaypool.com/page/ForthcomingSynthesisLectures}{Synthesis Digital Library} has many excellent short books/long tutorials on relevant topics
\begin{itemize}
\item \href{http://www.morganclaypool.com/toc/icr/1/1}{Information Concepts, Retrieval and Services}
\item \href{http://www.morganclaypool.com/toc/hlt/1/1}{Human Language Technology}
\item \href{http://www.morganclaypool.com/toc/aim/1/1}{Artificial Intelligence \& Machine Learning}
\end{itemize}

\item Journals: ACM TOIS, IRJ, IPM, ...
\item Conferences: SIGIR, CIKM, ECIR, WSDM, WWW, KDD, ACL, ...
\item \href{http://timan.cs.uiuc.edu/resources}{More info}
\item \href{http://searchuserinterfaces.com/}{Search User Interface}, by Marti Hearst, Cambridge University Press, 2009
\end{itemize}

\subsection{Main Techniques for Harnessing Big Text Data: Text Retrieval + Text Mining}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\linewidth]{big_text_data.png}
\end{figure}